### 2020-05-11 draft_analysis.Rmd
Merge predictions with the 2020 draft data Get the top 5 predictions Some reformatting

---
title: 'NFL Draft Analysis'
author: "Philip Ipe"
date: "5/8/2020"
output: html_document
knit: (function(input_file, encoding) {
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn = -1)
library(tidyverse)
library(rvest)
library(gridExtra)
setwd("/Users/PhilipIpe/Documents/Data Science/Projects/NFL Draft Analytics/")
```

# Part 1: Data Collection

We will scrape historical draft data from Pro-Football Reference from years 1970 till 2016. We chose to stop at 2016 since rookies usually sign 4 year deals, so the four year mark is typically when each team will finish evaulating each player they drafted, and make a decision on whether or not to resign the player. So, the jury is still out on players drafted after 2016.  We will filter out any selections after round 7, as since 1994, the NFL draft has only consisted of 7 rounds.

```{r scraping}

master_df <- data.frame()
for (year in 1970:2016){

  draft_url <- paste("https://www.pro-football-reference.com/years/", year, "/draft.htm", sep="")
  draft_html <- read_html(draft_url)

  df <- draft_html %>%
    html_node("table") %>%
    html_table()

  names(df) <- as.character(unlist(df[1,]))
  df <- df[-1,]

  # get rid of the columns we dont care about
  df <- df[,1:13]

  df$Rnd <- as.integer(df$Rnd)
  df$Age <- as.integer(df$Age)
  df$CarAV <- as.integer(df$CarAV)
  df$DrAV <- as.integer(df$DrAV)
  df$G <- as.integer(df$G)

  # filter out certain rounds
  df <- df %>%
    filter(Rnd < 8)

  # get rid of blank rows
  df <-df[!(df$Player=="Player"),]

  # add year drafted to each observation
  df <- df %>%
    mutate(year = year)

  master_df <- rbind(master_df, df)
}
```

We will also scrape the data from the 2020 draft for use in our machine learning problem and save it to our df variable.

```{r scrape2020}

draft_url <- "https://www.pro-football-reference.com/years/2020/draft.htm"
draft_html <- read_html(draft_url)

df_2020 <- draft_html %>%
  html_node("table") %>%
  html_table()

names(df_2020) <- as.character(unlist(df_2020[1,]))
df_2020 <- df_2020[-1,]
```

Next we will scrape the AP Offensive Rookie of the Year Winners from the same website.

```{r scrape_ROTY}

roty_url <- "https://www.pro-football-reference.com/awards/ap-offensive-rookie-of-the-year-award.htm"
draft_html <- read_html(roty_url)

roty_df <- draft_html %>%
  html_node("table") %>%
  html_table()
```

# Part 2: Data Processing and Preparation

Load the NFL draft history data, containing the draft data from rounds 1-7 from the 1970-2016 NFL draft. Remove all of the missing data entries, and convert the necessary columns to integers. 
```{r cars}
head(master_df)

master_df <- master_df[complete.cases(master_df[,12:13]),]
master_df$Rnd <- as.factor(master_df$Rnd)
master_df$Tm <- as.factor(master_df$Tm)
master_df$Pos <- as.factor(master_df$Pos)

master_df$Age <- as.integer(master_df$Age)
master_df$CarAV <- as.integer(master_df$CarAV)
master_df$DrAV <- as.integer(master_df$DrAV)
master_df$G <- as.integer(master_df$G)
master_df$Pick <- as.integer(master_df$Pick)
```

Now let's process our 2020 Draft data. First, we will get rid of columns we don't care about, then remove blank rows, add year drafted to each observation, and convert the necessary columns to integers, all like we did for the historical draft dataset. We will also filter the data to only include quarterbacks, wide receivers, and running backs.

```{r process2020}
# get rid of the columns we dont care about
df_2020 <- df_2020[,1:6]

# filter out certain rounds
df_2020 <- df_2020 %>%
  filter(Rnd < 8)

# get rid of blank rows
df_2020 <-df_2020[!(df_2020$Player=="Player"),]

# add year drafted to each observation, and filter out the defensive players.
df_2020 <- df_2020 %>%
  mutate(year = year) %>%
  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")

df_2020[is.na(df_2020)] <- 22

df_2020$Pick <- as.integer(df_2020$Pick)
df_2020$Age <- as.integer(df_2020$Age)
df_2020$Tm <- as.factor(df_2020$Tm)
df_2020$Pos <- as.factor(df_2020$Pos)
```

Now let's create a dataframe in which we can compare different positions that were drafted. First, we want to break up our data into 4 periods. Then for each period, we will take average career value of each position grouped by each round that the player was selected. We will make a new pos_df dataframe with these groupings.

```{r positions}
pos_df <- master_df %>%
  mutate(period = cut(year, breaks=4)) %>%
  group_by(period, Pos, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

head(pos_df)
```

Let's also add a dataframe that contains a variable that is TRUE when the player that was drafted was Offensive Rookie of the Year, and FALSE otherwise. We will also filter out all of the defensive players in this new dataframe.

```{r off_roty}
off_df <- master_df %>%
  mutate(roty = ifelse(Player %in% roty_df$Player, "Yes", "No")) %>%
  filter(Pos == "QB" | Pos == "WR" | Pos == "RB" | Pos == "G" | Pos == "T")

off_df$roty <- as.factor(off_df$roty)
```
# Part 3: Exploratory Data Analysis

## Trends in Average value
Has average value for players selected in different rounds gone up over time?
Have GMs gotten better at drafting in the first round, second round, etc...?

```{r avg_val_rd1}

avg_val_df <- master_df %>%
  group_by(year, Rnd) %>%
  summarise(Avg_AV = mean(CarAV))

avg_val_df %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=year, y=Avg_AV, color=Rnd)) +
    geom_point() +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for Players Drafted in Different Rounds Over Time") + 
    geom_smooth(method=lm) 
```

Seems like GMs have not gotten better at drafting in any of the first four rounds. In fact, we see a slight decline in average value for all rounds over time. We do see that GMs are selecting better players in the earlier rounds, as generally the average value for the earlier rounds is consistently higher than the lower rounds throughout the years. We see the trends spread out as time goes on as well.

## Analyzing the Quarterback position

Lets analyze the average values of different positions grouped by round and time period in which they were drafted. Let's start with the most important position in football.

```{r qb}
qb_bar <- pos_df %>%
  filter(Pos == "QB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for QBs Drafted in Different Rounds Over Different Time Periods")

qb_bar
```

In the past 20 years (2000-2020), it is much more likely to hit on a QB in the first round than in any of the other rounds. We see a huge drop in average value after the first round in the past 20 years. Average value dips in half after that, and continues to decrease until the 7th round. It is interesting to note that the 7th round averge value shoots up again to about the same average values of the 3rd and 4th round in the 2000s.

Before then, from 1990-2000, we see that the 6th round was a  value pick for QB, although Tom Brady might be the outlier that is skewing those numbers. But we also see that there was far more value in the second round QB selection from 1980-2000 than in the past 20 years.

Now let's focus our attention on the past 20 years and graph some visualizations of that data.
```{r qb2}

qb_av_pick <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=Pick, y=CarAV)) +
    geom_point()  +
    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for QBs Drafted in the 2000s") + 
    geom_smooth(method='loess')

qb_av_pick
```

As we would expect the trend is that there is a decline in value until about pick 100, where the career value seems to be about the same from about pick 100-250. This would suggest that seelcting a QB at picks 100-150 would, statistically speaking, provide less bang for your buck, as you could likely get a player of the same quality with some of your later draft capital.

```{r qb3}

qb_violin <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "QB") %>%
  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
    geom_violin()  +
    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for QBs Drafted in the 2000s") + 
    geom_smooth(method=lm)

qb_violin
```

As we would expect, the violin plots show us that most of the QBs selected in rounds 5-7 turn out to be busts, as most of the career values end up towards the bottom of the violin.

However, the distributions reaffirm what we have seen before in our Round 1 QBs. The spread of the data is far greater in round 1, and decreases heavily in round 2, and from there the decreases in spread are less drastic, suggesting that you have more of a shot at getting a good player in round 1 than in any other round.

## Analyzing the WR and RB positions

Next we will provide the same graphs to provide similar insight into the RB and WR selections of the draft.

It is interesting to note that the general trend has been to select RBs later in the draft, as RBs have always been considered as more expendable than a QB, and there seems to be a notion that you can always get a quality back in the later rounds, and that using first round capital on a RB is a poor management decision.

```{r rb_wr1}

rb_bar <- pos_df %>%
  filter(Pos == "RB") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for RBs Drafted in Different Rounds Over Different Time Periods")

wr_bar <- pos_df %>%
  filter(Pos == "WR") %>%
  group_by(Rnd) %>%
  ggplot(mapping=aes(x=period, y=Avg_AV, fill=Rnd)) +
    geom_bar(stat="identity", position = "dodge") +
    labs(x = "Year", y = "Average Career Value", title = "Average Value for WRs Drafted in Different Rounds Over Different Time Periods")

grid.arrange(qb_bar, rb_bar, wr_bar, ncol=3)
```

Not as drastic of a difference in rounds 1 and 2 in the WR group. The RB average value in the 2000s seems to resemble the QB position.

```{r rb_wr2}

rb_av_pick <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "RB") %>%
  ggplot(mapping=aes(x=Pick, y=CarAV)) +
    geom_point()  +
    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for RBs") + 
    geom_smooth(method='loess')

wr_av_pick <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "WR") %>%
  ggplot(mapping=aes(x=Pick, y=CarAV)) +
    geom_point()  +
    labs(x = "Pick", y = "Career Value", title = "Average Value over picks for WRs") + 
    geom_smooth(method='loess')

grid.arrange(qb_av_pick, rb_av_pick, wr_av_pick, ncol=3)

```

Much less steep of a drop off between round 1 to 2 in the WR group.

Again we see a trend that by pick 100 the value seems to hit a plateue.

```{r rb_wr3}

rb_violin <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "RB") %>%
  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
    geom_violin()  +
    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for RBs") + 
    geom_smooth(method=lm)

wr_violin <- master_df %>%
  filter(year >= 2000) %>%
  filter(Pos == "WR") %>%
  ggplot(mapping=aes(x=factor(Rnd), y=CarAV)) +
    geom_violin()  +
    labs(x = "Rnd", y = "Career Value", title = "Average Value over Round for WRs") + 
    geom_smooth(method=lm)

grid.arrange(qb_violin, rb_violin, wr_violin, ncol=3)
```

RBs in Round 1 always produce.
We again see that round 2 is a value round for WRs as it appears to resemble the type of value you get out of round 1.

# Part 4: Machine Learning

## Predicting the Rookie Offensive MVP

Now, right off the bat, we hope that the model will predict that Joe Burrow has a higher likelihood of winning the MVP.

First, lets control the resampling of our data. In this case, we’re going to cross-validate the data 3 times, therefore training it 3 times on different portions of the data before settling on the best tuning parameters (for gbm it is trees, shrinkage, and interaction depth).

```{r ml_control}

library(caret)

outcome_name <- 'roty'
predictor_names <- c('Tm', 'Pos', 'Pick')

objControl <- trainControl(method='cv', number=3, returnResamp='none', summaryFunction = twoClassSummary, classProbs = TRUE)

```

# Creating our Model
Now to actually train the model...

```{r train_model}

set.seed(123)
model <- train(off_df[,predictor_names], off_df[,outcome_name], 
               method = "gbm",
               trControl = objControl,
               metric = "ROC",
               preProc = c("center", "scale"))
```

```{r sum_model}
summary(model)
```

The summaries tell us how much influence each variable had on the model. It was interesting to see that position had such little influence in predicting the model. What is even more interesting is that the team the player was drafted to seemed to influence the model more than the pick number itself! Before I had trained the model, I expected pick, position, and then team to have relative influence in that order!

# Testing against 2017-2019 data

# Predicting the 2020 Rookie Offensive MVP

Now let's use the model to predict the probabilities that each player from the 2020 draft will become Offensive Rookie of the Year.  This code will give us the probability that the model is sure about its choice that the observation will fall under the "No" and "Yes" category.

```{r eval}

predictions <- predict(object=model, df_2020[,predictor_names], type='prob')
head(predictions)
```

Now let's join our prediction probabilities to the actual data, so we can look at the prediction probabilities for each player.

```{r join_probabilities}

df_2020 <- merge(df_2020, predictions, by = "row.names", all.x = FALSE, all.y = FALSE)
df_2020$No.x <- NULL
```

Now let's get the top 10 players that the model is most certain will be the Offensive Rookie of the Year

```{r top10_roty}

top5_roty <- df_2020 %>%
  top_n(5, Yes)

head(top5_roty)
```

Joe Burrow is our model's top prediction to win the MVP!!!